---
title: "Nutricast: Data preparation"
output: html_document
---

## Introduction

This script wrangles the various datasets used in the Nutricast app.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(countrycode) # fix and standardize country codes
library(here)
library(tidyverse)

data_dir <- here::here("data", "processed/")
if (dir.exists(data_dir) == F) {
  dir.create(data_dir, recursive = T)
} 

app_data_dir <- here::here("app", "data", "processed/")
if (dir.exists(app_data_dir) == F) {
  dir.create(app_data_dir, recursive = T)
} 
```

## Population Growth

Population data used in the app is from the World Bank. 

```{r}
# Population growth
national_pop_dat <- readRDS(here::here("data", "raw", "WB_UN_1960_2100_human_population_by_country.Rds"))

global_pop_dat <- readRDS(here::here("data", "raw", "WB_UN_1960_2100_human_population_global.Rds")) %>%
  mutate(country = "Global",
         iso3 = "Global")

pop_dat <- national_pop_dat %>%
  bind_rows(global_pop_dat) %>%
  dplyr::select(source, country, iso3, year, pop_size_05perc, pop_size_50perc, pop_size_95perc)

# Save 
saveRDS(pop_dat, paste0(data_dir, "01_01_population_growth_plot_data.Rds"))
saveRDS(pop_dat, paste0(app_data_dir, "01_01_population_growth_plot_data.Rds"))
```

## Nutrient Deficiencies

Not entirely sure where this data comes from

```{r, echo=FALSE}
# Load national nutrient deficit data
national_nutrient_deficiency_dat <- readRDS(here::here("data", "raw", "nutr_deficiencies_by_cntry_sex_age_2011.Rds")) %>%
  ungroup()

# The national level data is used in one plot as is - so let's save this for that
saveRDS(national_nutrient_deficiency_dat, paste0(data_dir, "02_03_nutritional_health_plot_data.Rds"))
saveRDS(national_nutrient_deficiency_dat, paste0(app_data_dir, "02_03_nutritional_health_plot_data.Rds"))

# Now let's wrangle it to combine with the global data for a different plot
national_nutrient_deficiency_dat_edit <- national_nutrient_deficiency_dat %>%
  dplyr::select(iso3, country, age, sex, nutrient, ndeficient, nhealthy)

global_nutrient_deficiency_dat <- readRDS(here::here("data", "raw", "nutr_deficiencies_by_sex_age_2011.Rds")) %>%
  ungroup() %>%
  mutate(iso3 = "Global", country = "Global") %>%
  dplyr::select(iso3, country, age, sex, nutrient, ndeficient, nhealthy) 

# Combine data for plot and format
nutrient_deficiency_dat <- national_nutrient_deficiency_dat_edit %>%
  bind_rows(global_nutrient_deficiency_dat) %>%
  filter(sex != "Children") %>% # Remove children (not symmetric)
  gather(key="type", value="npeople", 6:7) %>%  # Gather
  mutate(type=recode_factor(type,
                            "ndeficient"="Deficient",
                            "nhealthy"="Healthy"),
         sex=factor(sex, levels=c("Women", "Men"))) %>% 
  group_by(iso3, country, sex, age, nutrient) %>%
  mutate(ntotal = sum(npeople, na.rm = T)) %>%
  ungroup() %>%
  mutate(ppeople = (npeople/ntotal)*100) %>%
  mutate(npeople=ifelse(sex=="Men", npeople*-1, npeople),   # Make male values negative for plotting
         ppeople=ifelse(sex=="Men", ppeople*-1, ppeople))

# Save
saveRDS(nutrient_deficiency_dat, paste0(data_dir, "01_02a_nutrient_demand_plot_data.Rds"))
saveRDS(nutrient_deficiency_dat, paste0(app_data_dir, "01_02a_nutrient_demand_plot_data.Rds"))
```

## Projected Nutrient Demands

```{r}
national_nutrient_demand_dat <- readRDS(here::here("data", "raw", "1960_2100_nutrient_demand_by_country.Rds")) 
  # dplyr::select(-iso3) %>%
  # mutate(iso3 = countrycode(country, "country.name", "iso3c"))

global_nutrient_demand_dat <- readRDS(here::here("data", "raw", "1960_2100_nutrient_demand_global.Rds")) %>%
  mutate(country = "Global",
         iso3 = "Global")

nutrient_demand_dat <- national_nutrient_demand_dat %>%
  bind_rows(global_nutrient_demand_dat)

# Save
saveRDS(nutrient_demand_dat, paste0(data_dir, "01_02b_nutrient_demand_plot_data.Rds"))
saveRDS(nutrient_demand_dat, paste0(app_data_dir, "01_02b_nutrient_demand_plot_data.Rds"))
```

## Marine Capture & Mariculture Production

From the FAO.

```{r}
# Historical capture production by country (FAO)
national_capture_production_dat <- readRDS(here::here("data", "raw", "1950_2017_fao_landings_by_country_isscaap.Rds")) %>%
  dplyr::filter(!(isscaap %in% c("Freshwater molluscs", "Miscellaneous freshwater fishes", "Freshwater crustaceans", "River eels"))) %>% # remove freshwater production
  rename(country = country_use, iso3 = iso3_use) %>%
  mutate(plot_group = case_when(major_group == "Pisces" ~ "Finfish",
                                isscaap %in% c("Clams, cockles, arkshells", "Mussels", "Scallops, pectens", "Oysters") ~ "Bivalves",
                                TRUE ~ "Other")) %>%
  dplyr::filter(prod_type == "Landings")

# Use conversion factors to get edible meat equivalents. 
national_capture_production_dat <- national_capture_production_dat %>%
  rename(live_weight = prod_mt) %>%
  mutate(edible_meat = live_weight * conv_factor) %>%
  group_by(country, iso3, year, plot_group) %>%
  summarize(live_weight = sum(live_weight, na.rm = T),
            edible_meat = sum(edible_meat, na.rm = T)) %>%
  ungroup()

# Then join population data to get edible meat per capita. 
pop_dat_to_match <- pop_dat %>%
  group_by(iso3, year) %>%
  summarize(capita = unique(pop_size_50perc)) %>%
  ungroup()

national_capture_production_dat <- national_capture_production_dat %>%
  left_join(pop_dat_to_match, by = c("iso3", "year")) %>%
  mutate(edible_meat_per_capita = edible_meat/capita) %>%
  mutate(source = "Marine Capture")

# Now do global
global_capture_production_dat <- national_capture_production_dat %>%
  group_by(year, plot_group, source) %>%
  summarize(live_weight = sum(live_weight, na.rm = T),
            edible_meat = sum(edible_meat, na.rm = T),
            capita = sum(capita, na.rm = T)) %>%
  ungroup() %>%
  mutate(edible_meat_per_capita = edible_meat/capita) %>%
  mutate(country = "Global", iso3 = "Global")

### Now do aquaculture production -----------------

# Historical aquaculture production by country (FAO)
national_aquaculture_production_dat <- readRDS(here::here("data", "raw", "1950_2017_fao_aquaculture_data.Rds")) %>%
  dplyr::filter(environment != "Freshwater") %>% # remove freshwater production
  mutate(plot_group = case_when(major_group == "Pisces" ~ "Finfish",
                                order == "Bivalvia" ~ "Bivalves",
                                TRUE ~ "Other")) 

# Load conversion factors to apply live weight to edible meat conversions
lw_to_em_conversion <- read_csv(here::here("data", "isscaap_conversion_factors.csv")) %>%
  dplyr::select(isscaap, conv_factor = catch2meat)

national_aquaculture_production_dat <- national_aquaculture_production_dat %>%
  left_join(lw_to_em_conversion, by = c("isscaap")) %>%
  rename(country = country_orig, iso3 = iso3_orig, live_weight = quantity_mt) %>%
  mutate(edible_meat = live_weight * conv_factor) %>%
  group_by(country, iso3, year, plot_group) %>%
  summarize(live_weight = sum(live_weight, na.rm = T),
            edible_meat = sum(edible_meat, na.rm = T)) %>%
  ungroup()

# Now apply population numbers
national_aquaculture_production_dat <- national_aquaculture_production_dat %>%
  left_join(pop_dat_to_match, by = c("iso3", "year")) %>%
  mutate(edible_meat_per_capita = edible_meat/capita) %>%
  mutate(source = "Mariculture")

# Now do global
global_aquaculture_production_dat <- national_aquaculture_production_dat %>%
  group_by(year, plot_group, source) %>%
  summarize(live_weight = sum(live_weight, na.rm = T),
            edible_meat = sum(edible_meat, na.rm = T),
            capita = sum(capita, na.rm = T)) %>%
  ungroup() %>%
  mutate(edible_meat_per_capita = edible_meat/capita) %>%
  mutate(country = "Global", iso3 = "Global")

# Output plot dat
production_plot_dat <- national_capture_production_dat %>%
  bind_rows(global_capture_production_dat) %>%
  bind_rows(national_aquaculture_production_dat) %>%
  bind_rows(global_aquaculture_production_dat) %>%
  dplyr::filter(year >= 1960)

# Save
saveRDS(production_plot_dat, paste0(data_dir, "01_03a_fisheries_mariculture_production_plot_data.Rds"))
saveRDS(production_plot_dat, paste0(app_data_dir, "01_03a_fisheries_mariculture_production_plot_data.Rds"))
```

## Nutrients from Seafood and Seafood as a Part of the Diet

This is from the Genus project. 

```{r}
national_nutrient_from_seafood_dat <- readRDS(here::here("data", "raw", "genus_pnutrient_seafood_by_cntry_2011.Rds"))

national_diet_from_seafood_dat <- readRDS(here::here("data", "raw", "genus_pdiet_seafood_by_cntry_year.Rds"))
  
protein_dat <- national_nutrient_from_seafood_dat %>% 
  filter(nutrient=="Protein") %>% 
  mutate(prop_seafood_cap = pmin(prop_seafood, 0.1))

# Save
saveRDS(protein_dat, paste0(data_dir, "01_03b_protein_from_seafood_plot_data.Rds"))
saveRDS(protein_dat, paste0(app_data_dir, "01_03b_protein_from_seafood_plot_data.Rds"))

# Save
saveRDS(national_diet_from_seafood_dat, paste0(data_dir, "01_03c_seafood_consumption_plot_data_diet.Rds"))
saveRDS(national_diet_from_seafood_dat, paste0(app_data_dir, "01_03c_seafood_consumption_plot_data_diet.Rds"))

# Save
saveRDS(national_nutrient_from_seafood_dat, paste0(data_dir, "01_03c_seafood_consumption_plot_data_nutrients.Rds"))
saveRDS(national_nutrient_from_seafood_dat, paste0(app_data_dir, "01_03c_seafood_consumption_plot_data_nutrients.Rds"))
```



