---
title: "Nutricast - Process Species Suitability Rasters"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(countrycode) # fix and standardize country codes
library(here)
library(tidyverse)
library(raster)
library(stringr)
library(leaflet)
library(rgdal)

data_dir <- here::here("data", "processed", "species_rasters/")
if (dir.exists(data_dir) == F) {
  dir.create(data_dir, recursive = T)
} 

app_data_dir <- here::here("app", "data", "processed", "species_rasters/")
if (dir.exists(app_data_dir) == F) {
  dir.create(app_data_dir, recursive = T)
} 
```

## Testing

Let's start by reading in a sample file to see how the new files (from August 2020) have changed relative to the old ones. 

```{r cars}
# Get our files of interest... 
raster_data_path <- "/Volumes/GoogleDrive/Shared\ drives/emlab/projects/current-projects/blue-paper-2/data/output/raw"
raster_files <- list.files(path = raster_data_path, full.names = T, pattern = "*.Rds")

# Practice extracting the data we need to extract from a single file... 
file_name <- str_replace(raster_files[1], paste0(raster_data_path, "/"), "")
scenario_name <- str_extract(file_name, "[^_]+")
genus_name <- str_extract(str_replace(file_name, paste0(scenario_name, "_"), ""), "[^_]+")
species_name <- str_replace(str_extract(str_replace(file_name, paste0(scenario_name, "_", genus_name, "_"), ""), "[^_]+"), ".Rds", "")
full_species_name <- paste(genus_name, species_name)

# Now load test file... 
test_file <- readRDS(raster_files[1]) %>%
  dplyr::select(x, y, viable)

head(test_file)

# Let's try the practice of turning it into a raster and adding to leaflet... we'll then compare to the data frame... 
coordinates(test_file) <- ~ x + y
gridded(test_file) <- TRUE
crs(test_file) <- sp::CRS("+proj=moll +ellps=WGS84 +datum=WGS84 +no_defs")
raster_file <- raster(test_file)

# Now add to leaflet - seems to work pretty well. 
pal <- colorNumeric(c("green"), values(raster_file),
  na.color = "transparent")

leaflet() %>% addTiles() %>%
  addRasterImage(raster_file, colors = pal, opacity = 0.8) %>%
  addLegend(pal = pal, values = values(raster_file),
    title = "Suitability")

# Now let's save, and reread and see how large this file is/how long it takes. 
saveRDS(raster_file, paste0(data_dir, "test_file.Rds"))
```

Okie, this seems to be working well... Let's see if we can now make raster stacks or bricks by species. 

```{r}
# Respecify our files of interest... 
raster_data_path <- "/Volumes/GoogleDrive/Shared\ drives/emlab/projects/current-projects/blue-paper-2/data/output/raw"
raster_files <- list.files(path = raster_data_path, full.names = T, pattern = "*.Rds")

# Now lets see if we can get a list of all species... Practice extracting the data we need to extract from all files. Oh good they vectorized these. 
file_names_short <- str_replace(raster_files, paste0(raster_data_path, "/"), "")
scenario_names <- str_extract(file_names_short, "[^_]+")
genus_names <- str_extract(str_replace(file_names_short, paste0(scenario_names, "_"), ""), "[^_]+")
species_names <- str_replace(str_extract(str_replace(file_names_short, paste0(scenario_names, "_", genus_names, "_"), ""), "[^_]+"), ".Rds", "")
full_species_names <- paste(genus_names, species_names)

# Ok let's see how many species we have. Looks like 193. 
unique_scenario_names <- unique(scenario_names)
unique_species_names <- unique(full_species_names)
```

```{r}
# Testing values
file_dir <- raster_data_path
all_file_names <- raster_files
all_scenario_names <- unique_scenario_names
species_name <- unique_species_names[1]

# Now let's set up a function for reading in all files cooresponding to a single species. 
ReadAllSpeciesFiles <- function(file_dir, all_file_names, all_scenario_names, species_name){
  
  # Format the species name, and generate all possible files for this species. We do not know if all of these files will actually exist for each species. 
  species_name_separated <- str_replace(species_name, " ", "_")
  possible_file_names <- paste0(file_dir, "/", all_scenario_names, "_", species_name_separated, ".Rds")
  
  # Find the matching files
  matching_files <- all_file_names[all_file_names %in% possible_file_names]
  
  # Get the scenario names for matching files (important later for naming the layers in the raster object)
  matching_files_short <- str_replace(matching_files, paste0(file_dir, "/"), "")
  scenario_names <- str_extract(matching_files_short, "[^_]+")
  
  # Now load first file and turn into a raster
  file_1 <- readRDS(matching_files[1]) %>%
    dplyr::filter(period == "2091-2100")%>%
    dplyr::select(x, y, viable)
  
  coordinates(file_1) <- ~ x + y
  gridded(file_1) <- TRUE
  crs(file_1) <- sp::CRS("+proj=moll +ellps=WGS84 +datum=WGS84 +no_defs")
  raster_layer_1 <- raster(file_1)
  
  tibble_out <- tibble(species = species_name, scenario = scenario_names[1], suitability = list(raster_layer_1))
  
  # Now load other files (if applicable) and add into our tibble
  for(i in 2:length(matching_files)){
    
    # Now load first file and turn into a raster
    file_add <- readRDS(matching_files[i]) %>%
      dplyr::filter(period == "2091-2100") %>%
      dplyr::select(x, y, viable)
  
    coordinates(file_add) <- ~ x + y
    gridded(file_add) <- TRUE
    crs(file_add) <- sp::CRS("+proj=moll +ellps=WGS84 +datum=WGS84 +no_defs")
    raster_layer_add <- raster(file_add)
    
    tibble_entry_add <- tibble(species = species_name, scenario = scenario_names[i], suitability = list(raster_layer_add))
    
    # Stack
    tibble_out <- tibble_out %>%
      bind_rows(tibble_entry_add)
  }

  saveRDS(tibble_out, paste0(data_dir, species_name_separated, ".Rds"))
  
  print(paste0(species_name, " Completed"))
  
  return(tibble_out)
  
}
```

## Merge data files by species - Practice run

```{r}
# Run function for one species and see if it works
test_species <- ReadAllSpeciesFiles(file_dir, all_file_names, all_scenario_names, species_name)

# Practice extracting plot data
plot_dat <- test_species %>%
  dplyr::filter(scenario == "RCP85")

raster_plot <- unlist(plot_dat$suitability)[[1]]

# Now add to leaflet
pal <- colorNumeric(c("green"), values(raster_plot),
  na.color = "transparent")

leaflet() %>% addTiles() %>%
  addRasterImage(raster_plot, colors = pal, opacity = 0.8) %>%
  addLegend(pal = pal, values = values(raster_plot),
    title = "Suitability")
```

I'm starting to think that I might be able to put all species rasters in one tibble... This would speed things up immensely in only having to load in one file at the beginning.. Let's try it. 

```{r}
output_tibble <- NULL
files_to_run <- unique_species_names[!(unique_species_names %in% c("Anguilla anguilla (1)", "Coryphaena hippurus (1)", "Dicentrarchus labrax (1)", "Pollachius pollachius (1)"))]

for(p in 1:length(unique_species_names)){
  
  species_to_add_to_tibble <- ReadAllSpeciesFiles(file_dir, all_file_names, all_scenario_names, files_to_run[p])
  
  output_tibble <- output_tibble %>%
    bind_rows(species_to_add_to_tibble)
}
```

