---
title: "Nutricast - Process Species Suitability Rasters"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(countrycode) # fix and standardize country codes
library(here)
library(tidyverse)
library(raster)
library(stringr)
library(leaflet)
library(rgdal)

data_dir <- here::here("data", "processed", "species_rasters/")
if (dir.exists(data_dir) == F) {
  dir.create(data_dir, recursive = T)
} 

app_data_dir <- here::here("app", "data", "processed", "species_rasters/")
if (dir.exists(app_data_dir) == F) {
  dir.create(app_data_dir, recursive = T)
} 
```

## Testing

Let's start by reading in a sample file to see how the new files (from August 2020) have changed relative to the old ones. 

```{r cars}
# # Get our files of interest... 
# raster_data_path <- "/Volumes/GoogleDrive/Shared\ drives/emlab/projects/current-projects/blue-paper-2/data/output/raw"
# raster_files <- list.files(path = raster_data_path, full.names = T, pattern = "*.Rds")
# 
# # Practice extracting the data we need to extract from a single file... 
# file_name <- str_replace(raster_files[1], paste0(raster_data_path, "/"), "")
# scenario_name <- str_extract(file_name, "[^_]+")
# genus_name <- str_extract(str_replace(file_name, paste0(scenario_name, "_"), ""), "[^_]+")
# species_name <- str_replace(str_extract(str_replace(file_name, paste0(scenario_name, "_", genus_name, "_"), ""), "[^_]+"), ".Rds", "")
# full_species_name <- paste(genus_name, species_name)
# 
# # Now load test file... 
# test_file <- readRDS(raster_files[1]) %>%
#   dplyr::select(x, y, viable)
# 
# head(test_file)
# 
# # Let's try the practice of turning it into a raster and adding to leaflet... we'll then compare to the data frame... 
# coordinates(test_file) <- ~ x + y
# gridded(test_file) <- TRUE
# crs(test_file) <- sp::CRS("+proj=moll +ellps=WGS84 +datum=WGS84 +no_defs")
# raster_file <- raster(test_file)
# 
# # Now add to leaflet - seems to work pretty well. 
# pal <- colorNumeric(c("green"), values(raster_file),
#   na.color = "transparent")
# 
# leaflet() %>% addTiles() %>%
#   addRasterImage(raster_file, colors = pal, opacity = 0.8)
# 
# # Now let's save, and reread and see how large this file is/how long it takes. 
# saveRDS(raster_file, paste0(data_dir, "test_file.Rds"))
```

Okie, this seems to be working well... Let's see if we can now make raster stacks or bricks by species. 

```{r}
# Respecify our files of interest... 
raster_data_path <- "/Volumes/GoogleDrive/Shared\ drives/emlab/projects/current-projects/blue-paper-2/data/output/raw"
raster_files <- list.files(path = raster_data_path, full.names = T, pattern = "*.Rds")

# Now lets see if we can get a list of all species... Practice extracting the data we need to extract from all files. Oh good they vectorized these. 
file_names_short <- str_replace(raster_files, paste0(raster_data_path, "/"), "")
scenario_names <- str_extract(file_names_short, "[^_]+")
genus_names <- str_extract(str_replace(file_names_short, paste0(scenario_names, "_"), ""), "[^_]+")
species_names <- str_replace(str_extract(str_replace(file_names_short, paste0(scenario_names, "_", genus_names, "_"), ""), "[^_]+"), ".Rds", "")
full_species_names <- paste(genus_names, species_names)

# Ok let's see how many species we have. Looks like 193. 
unique_scenario_names <- unique(scenario_names)
unique_species_names <- unique(full_species_names)
```

```{r}
#  Values
file_dir <- raster_data_path
all_file_names <- raster_files
all_scenario_names <- unique_scenario_names

# Now let's set up a function for reading in all files cooresponding to a single species. 
ReadAllSpeciesFiles <- function(file_dir, all_file_names, all_scenario_names, species_name){
  
  # Format the species name, and generate all possible files for this species. We do not know if all of these files will actually exist for each species. 
  species_name_separated <- str_replace(species_name, " ", "_")
  possible_file_names <- paste0(file_dir, "/", all_scenario_names, "_", species_name_separated, ".Rds")
  
  # Find the matching files
  matching_files <- all_file_names[all_file_names %in% possible_file_names]
  
  # Get the scenario names for matching files (important later for naming the layers in the raster object)
  matching_files_short <- str_replace(matching_files, paste0(file_dir, "/"), "")
  scenario_names <- str_extract(matching_files_short, "[^_]+")
  
  # Now load first file and turn into a raster
  file_1 <- readRDS(matching_files[1]) %>%
    dplyr::filter(period == "2091-2100")%>%
    dplyr::select(x, y, viable)
  
  coordinates(file_1) <- ~ x + y
  gridded(file_1) <- TRUE
  crs(file_1) <- sp::CRS("+proj=moll +ellps=WGS84 +datum=WGS84 +no_defs")
  raster_layer_1 <- raster(file_1)
  
  tibble_out <- tibble(species = species_name, scenario = scenario_names[1], suitability = list(raster_layer_1))
  
  # Now load other files (if applicable) and add into our tibble
  for(i in 2:length(matching_files)){
    
    # Now load first file and turn into a raster
    file_add <- readRDS(matching_files[i]) %>%
      dplyr::filter(period == "2091-2100") %>%
      dplyr::select(x, y, viable)
  
    coordinates(file_add) <- ~ x + y
    gridded(file_add) <- TRUE
    crs(file_add) <- sp::CRS("+proj=moll +ellps=WGS84 +datum=WGS84 +no_defs")
    raster_layer_add <- raster(file_add)
    
    tibble_entry_add <- tibble(species = species_name, scenario = scenario_names[i], suitability = list(raster_layer_add))
    
    # Stack
    tibble_out <- tibble_out %>%
      bind_rows(tibble_entry_add)
  }

  saveRDS(tibble_out, paste0(data_dir, species_name_separated, ".Rds"))
  
  print(paste0(species_name, " Completed"))
  
  return(tibble_out)
  
}
```

## Merge data files by species - Practice run

```{r}
# # Run function for one species and see if it works
# test_species <- ReadAllSpeciesFiles(file_dir, all_file_names, all_scenario_names, species_name)
# 
# # Practice extracting plot data
# plot_dat <- test_species %>%
#   dplyr::filter(scenario == "RCP85")
# 
# raster_plot <- unlist(plot_dat$suitability)[[1]]
# 
# # Now add to leaflet
# pal <- colorNumeric(c("green"), values(raster_plot),
#   na.color = "transparent")
# 
# leaflet() %>% addTiles() %>%
#   addRasterImage(raster_plot, colors = pal, opacity = 0.8) %>%
#   addLegend(pal = pal, values = values(raster_plot),
#     title = "Suitability")
```

I'm starting to think that I might be able to put all species rasters in one tibble... This would speed things up immensely in only having to load in one file at the beginning.. Let's try it. 

```{r}
problem_species <- c("Anguilla anguilla (1)", "Coryphaena hippurus (1)", "Dicentrarchus labrax (1)", "Pollachius pollachius (1)", "Argopecten purpuratus", "Cerastoderma edule", "Mya arenaria", "Mytilus chilensis", "Mytilus platensis", "Ostrea chilensis", "Ostrea edulis", "Pecten jacobaeus", "Pecten maximus", "Saxidomus gigantea", "Scrobicularia plana", "Thunnus orientalis", "Venerupis rhomboides")

output_tibble <- NULL
files_to_run <- unique_species_names[!(unique_species_names %in% problem_species)]

for(p in 1:length(files_to_run)){
  
  species_to_add_to_tibble <- ReadAllSpeciesFiles(file_dir, all_file_names, all_scenario_names, files_to_run[p])
  
  output_tibble <- output_tibble %>%
    bind_rows(species_to_add_to_tibble)
}

saveRDS(output_tibble, paste0(data_dir, "species_rasters_tibble.Rds"))
saveRDS(output_tibble, paste0(app_data_dir, "species_rasters_tibble.Rds"))
```

## Create lookup files linking species mariculture projections to eezs

```{r}
### Section 1) Get EEZ codes cooresponding to each x/y position
eez_raster <- raster::raster(here::here("data", "eezs_v10_raster_10km.tif"))

eez_info <- raster::as.data.frame(eez_raster, xy=T) %>%
  mutate(x = round(x, 0),
         y = round(y, 0)) %>%
  rename(eez_id = eezs_v10_raster_10km) %>%
  mutate(x_y_id = paste0(x, "/", y)) %>%
  dplyr::select(x_y_id, eez_id)

# Now define new data directories
data_dir <- here::here("data", "processed/")
if (dir.exists(data_dir) == F) {
  dir.create(data_dir, recursive = T)
} 

app_data_dir <- here::here("app", "data", "processed/")
if (dir.exists(app_data_dir) == F) {
  dir.create(app_data_dir, recursive = T)
} 
```

We'll start by matching data files from RCP 2.6

```{r}
rcp26_dat_files <- list.files(path = raster_data_path,
                              pattern = "RCP26",
                              full.names = T)

# Make blank data frame
rcp_26_dat_all <- data.frame()

for(i in 1:length(rcp26_dat_files)){
  
  text <- str_split(str_replace(rcp26_dat_files[i], ".Rds", ""), "_")
  species <- paste(text[[1]][2], text[[1]][3])
  
  if(species %in% problem_species){
    
    rcp_26_dat_all <- rcp_26_dat_all

  }else{
    
    foo <- readRDS(rcp26_dat_files[i]) %>%
      mutate(x = round(x, 0),
           y = round(y, 0)) %>%
      mutate(x_y_id = paste0(x, "/", y)) %>%
      left_join(eez_info, by = "x_y_id") %>%
      group_by(eez_id, period, viable) %>%
      summarize(n_points = n_distinct(x_y_id),
              prod_mt_yr = sum(prod_mt_yr, na.rm = T),
              revenue_usd_yr = sum(revenue_usd_yr, na.rm = T),
              cost_usd_yr = sum(cost_usd_yr, na.rm = T),
              profits_usd_yr = sum(profits_usd_yr, na.rm = T)) %>%
      ungroup() %>%
      mutate(species = species)
  
    rcp_26_dat_all <- rcp_26_dat_all %>%
      bind_rows(foo)
  }

}

write_csv(rcp_26_dat_all, paste0(data_dir, "rcp_26_species_dat_by_eez.csv"))
write_csv(rcp_26_dat_all, paste0(app_data_dir, "rcp_26_species_dat_by_eez.csv"))

```

Now RCP 4.5

```{r}
rcp45_dat_files <- list.files(path = raster_data_path,
                              pattern = "RCP45",
                              full.names = T)

# Make blank data frame
rcp_45_dat_all <- data.frame()

for(i in 1:length(rcp45_dat_files)){
  
  text <- str_split(str_replace(rcp45_dat_files[i], ".Rds", ""), "_")
  species <- paste(text[[1]][2], text[[1]][3])
  
  if(species %in% problem_species){
    
    rcp_45_dat_all <- rcp_45_dat_all

  }else{
    
    foo <- readRDS(rcp45_dat_files[i]) %>%
      mutate(x = round(x, 0),
           y = round(y, 0)) %>%
      mutate(x_y_id = paste0(x, "/", y)) %>%
      left_join(eez_info, by = "x_y_id") %>%
      group_by(eez_id, period, viable) %>%
      summarize(n_points = n_distinct(x_y_id),
              prod_mt_yr = sum(prod_mt_yr, na.rm = T),
              revenue_usd_yr = sum(revenue_usd_yr, na.rm = T),
              cost_usd_yr = sum(cost_usd_yr, na.rm = T),
              profits_usd_yr = sum(profits_usd_yr, na.rm = T)) %>%
      ungroup() %>%
      mutate(species = species)
  
    rcp_45_dat_all <- rcp_45_dat_all %>%
      bind_rows(foo)
  }

}

write_csv(rcp_45_dat_all, paste0(data_dir, "rcp_45_species_dat_by_eez.csv"))
write_csv(rcp_45_dat_all, paste0(app_data_dir, "rcp_45_species_dat_by_eez.csv"))

```

Now RCP 6.0

```{r}
rcp60_dat_files <- list.files(path = raster_data_path,
                              pattern = "RCP60",
                              full.names = T)

# Make blank data frame
rcp_60_dat_all <- data.frame()

for(i in 1:length(rcp60_dat_files)){
  
  text <- str_split(str_replace(rcp60_dat_files[i], ".Rds", ""), "_")
  species <- paste(text[[1]][2], text[[1]][3])
  
  if(species %in% problem_species){
    
    rcp_60_dat_all <- rcp_60_dat_all

  }else{
    
    foo <- readRDS(rcp60_dat_files[i]) %>%
      mutate(x = round(x, 0),
           y = round(y, 0)) %>%
      mutate(x_y_id = paste0(x, "/", y)) %>%
      left_join(eez_info, by = "x_y_id") %>%
      group_by(eez_id, period, viable) %>%
      summarize(n_points = n_distinct(x_y_id),
              prod_mt_yr = sum(prod_mt_yr, na.rm = T),
              revenue_usd_yr = sum(revenue_usd_yr, na.rm = T),
              cost_usd_yr = sum(cost_usd_yr, na.rm = T),
              profits_usd_yr = sum(profits_usd_yr, na.rm = T)) %>%
      ungroup() %>%
      mutate(species = species)
  
    rcp_60_dat_all <- rcp_60_dat_all %>%
      bind_rows(foo)
  }

}

write_csv(rcp_60_dat_all, paste0(data_dir, "rcp_60_species_dat_by_eez.csv"))
write_csv(rcp_60_dat_all, paste0(app_data_dir, "rcp_60_species_dat_by_eez.csv"))

```

Last RCP 8.5

```{r}
rcp85_dat_files <- list.files(path = raster_data_path,
                              pattern = "RCP85",
                              full.names = T)

# Make blank data frame
rcp_85_dat_all <- data.frame()

for(i in 1:length(rcp85_dat_files)){
  
  text <- str_split(str_replace(rcp85_dat_files[i], ".Rds", ""), "_")
  species <- paste(text[[1]][2], text[[1]][3])
  
  if(species %in% problem_species){
    
    rcp_85_dat_all <- rcp_85_dat_all

  }else{
    
    foo <- readRDS(rcp85_dat_files[i]) %>%
      mutate(x = round(x, 0),
           y = round(y, 0)) %>%
      mutate(x_y_id = paste0(x, "/", y)) %>%
      left_join(eez_info, by = "x_y_id") %>%
      group_by(eez_id, period, viable) %>%
      summarize(n_points = n_distinct(x_y_id),
              prod_mt_yr = sum(prod_mt_yr, na.rm = T),
              revenue_usd_yr = sum(revenue_usd_yr, na.rm = T),
              cost_usd_yr = sum(cost_usd_yr, na.rm = T),
              profits_usd_yr = sum(profits_usd_yr, na.rm = T)) %>%
      ungroup() %>%
      mutate(species = species)
  
    rcp_85_dat_all <- rcp_85_dat_all %>%
      bind_rows(foo)
  }

}

write_csv(rcp_85_dat_all, paste0(data_dir, "rcp_85_species_dat_by_eez.csv"))
write_csv(rcp_85_dat_all, paste0(app_data_dir, "rcp_85_species_dat_by_eez.csv"))

```

```{r}
rcp_26_projections <- rcp_26_dat_all %>%
  mutate(scenario = "RCP 2.6")

rcp_45_projections <- rcp_45_dat_all %>%
  mutate(scenario = "RCP 4.5")

rcp_60_projections <- rcp_60_dat_all %>%
  mutate(scenario = "RCP 6.0")

rcp_85_projections <- rcp_85_dat_all %>%
  mutate(scenario = "RCP 8.5")

rcp_projections <- rcp_26_projections %>%
  bind_rows(rcp_45_projections) %>%
  bind_rows(rcp_60_projections) %>%
  bind_rows(rcp_85_projections)

write_csv(rcp_projections, paste0(data_dir, "rcp_all_dat_by_eez.csv"))
write_csv(rcp_projections, paste0(app_data_dir, "rcp_all_dat_by_eez.csv"))
```

