---
title: "Nutricast - Shapefile processing"
output: html_document
---

## Introduction

This script modifies v10 of the EEZ shapefile for use in our mariculture explorer plot. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(tidyverse)

data_dir <- here::here("data", "processed", "shapefiles/")
if (dir.exists(data_dir) == F) {
  dir.create(data_dir, recursive = T)
} 

app_data_dir <- here::here("app", "data", "processed", "shapefiles/")
if (dir.exists(app_data_dir) == F) {
  dir.create(app_data_dir, recursive = T)
} 
```

## Shapefile

For simplicity, we're going to use a simplified version of the v10 Marine Regions shapefile, which has FAO regions (on the high seas) added. 

```{r cars}
# Load data
eez_info <- read_csv(here::here("data", "eez_info_table.csv"))
shp <- read_sf(here::here("data", "shapefiles", "eez_v10_fao_combined_simple"), layer = "eez_v10_fao_combined_simple")

# Filter out FAO regions
shp <- shp %>%
  dplyr::filter(is.na(zone)) %>%
  dplyr::select(eez_id = mrgid, name, type, ter1_iso3 = iso_tr1) %>%
  left_join(eez_info %>% dplyr::select(eez_id, sovereign1_iso3), by = "eez_id") %>%
  mutate(sovereign1_iso3 = ifelse(is.na(sovereign1_iso3), ter1_iso3, sovereign1_iso3))

write_sf(shp, paste0(app_data_dir, "eez_v10_simple"), "eez_v10_simple", delete_layer = T, driver = "ESRI Shapefile")

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
